Models of gravitational lens candidates from Space Warps CFHTLS

### 2017-09-07 parts/morph.tex
make fig uniform
my hands are typing words

The main input of a modeller to the process is a markup of the
candidate lens system, which we call a spaghetti diagram.  This is a
sketch of the arrival-time surface from a point-like source, with
proposed locations of maxima, minima and saddle points, and an implied
time-ordering of the images.  Such information encodes a starting
proposal for the mass distribution.  A spaghetti diagram is thus a
completely abstract construction, and moreover it refers to a
simplified system with a point source.  However, spaghetti diagrams
are intuitive because they tend to resemble the form of lensed arcs
\citep[see fig.3 of][]{2008MNRAS.383..857F}, and of course they are
simple to draw, and easy to vary and refine in an open collaborative
environment. This makes them very practical for non-professional lens
enthusiasts in the citizen-science community.  Details and tests are
given in \citet{2015MNRAS.447.2170K}.

@@

The fourth diagnostic is based on the fact that the arrangement of
lensed images of a pointlike source through a non-circular
gravitational lens depends on the location of the source relative to
the long and short axes of the lens.  This dependence is quite robust
and independent of many other details of the lensing mass
distribution.  Sources close to being dead-centre behind a lens tend
to produce quads; sources at larger transverse distance tend to
produce doubles.  We denote these as Q and D, respectively, and add a
prefix to the Q systems, as follows: we write LQ if the source is
inferred as displaced along the long axis of the lens, SQ if displaced
along the short axis, IQ if inclined to both axes, CQ if only very little or
no displacement is evident.  Although the
unlensed source and its location are obviously not seen, the LQ, SQ, IQ and CQ
cases correspond to easily-seen image morphologies \citep[see,
  e.g.,][]{2003AJ....125.2769S}.
\begin{itemize}
\item The simplest is the SQ case: this creates a saddle point and two
  minima in an arc, with the second saddle point on the other side of
  the lensing galaxy, closer to the galaxy than the arc.  \sw{58}
  and \sw{28} in the upper row of Fig.~\ref{fig:splinput} are typical
  examples of SQ.  If the source were moved outwards along the short
  axis, the minimum-saddle-minimum set would merge into a single minimum,
  leaving a D system; the transition is known as a cusp catastrophe.
\item The middle row of Fig.~\ref{fig:splinput} shows three IQ
  systems: \sw{05}, \sw{42} and \sw{19}.  This type has a characteristic
  asymmetry, often with two images close together.  If the source
  were moved outwards, the minimum-saddle pair would merge and
  mutually cancel, leaving again a D system. This transition is known
  as a fold catastrophe.
\item The lower row of Fig.~\ref{fig:splinput} shows three LQ
  systems: \sw{09}, \sw{29} and \sw{02}. The failed model, \sw{57}, at top right
  may also belong to this category.  Here the images have a
  fairly symmetric arrangement with an arc and a counter-image on the
  other side, but the spaghetti diagram is completely asymmetric.  If
  the source were moved outwards along the long axis, the
  saddle-minimum-saddle set  would merge into a single saddle --- another
  form of cusp catastrophe.  LQ can be visually distinguished from SQ
  by the relative distances of the arc and the counter-image.  For
  LQ the arc is closer, for SQ the counter-image is closer.
\end{itemize}
CQ systems are often called `cross' or `Einstein-cross' systems; IQ
systems are sometimes called `folds', with `cusp' commonly used for
both SQ and LQ.  The labels `short-axis quad' and so on are not
standard in the literature, but the morphological classification they
express is familiar to experienced modellers.  Hence they can be
useful to researchers wishing to apply other modelling methods to the
same systems.

>>>>

The main input of a modeller to the process is a markup of the
candidate lens system, which we call a spaghetti diagram.  This is a
sketch of the arrival-time surface from a point-like source, with
proposed locations of maxima, minima and saddle points, and an implied
time-ordering of the images.  Such information encodes a starting
proposal for the mass distribution.  A spaghetti diagram is thus a
completely abstract construction, and moreover it refers to a
simplified system with a point source.  However, spaghetti diagrams
are intuitive because they tend to resemble the form of lensed arcs
\citep[see Fig.3 of][]{2008MNRAS.383..857F}, and of course they are
simple to draw, and easy to vary and refine in an open collaborative
environment. This makes them very practical for non-professional lens
enthusiasts in the citizen-science community.  Details and tests are
given in \citet{2015MNRAS.447.2170K}.

@@

The fourth diagnostic is based on the fact that the arrangement of
lensed images of a pointlike source through a non-circular
gravitational lens depends on the location of the source relative to
the long and short axes of the lens.  This dependence is quite robust
and independent of many other details of the lensing mass
distribution.  Sources close to being dead-centre behind a lens tend
to produce quads; sources at larger transverse distance tend to
produce doubles.  We denote these as Q and D, respectively, and add a
prefix to the Q systems, as follows: we write LQ if the source is
inferred as displaced along the long axis of the lens, SQ if displaced
along the short axis, IQ if inclined to both axes, CQ if only very little or
no displacement is evident.  Although the
unlensed source and its location are obviously not seen, the LQ, SQ, IQ and CQ
cases correspond to easily-seen image morphologies \citep[see,
  e.g.,][]{2003AJ....125.2769S}.
\begin{itemize}
\item The simplest is the LQ case: this creates a saddle point and two
  minima in an arc, with the second saddle point on the other side of
  the lensing galaxy, closer to the galaxy than the arc.  \sw{58}
  and \sw{28} in the upper row of Fig.~\ref{fig:splinput} are typical
  examples of LQ.  If the source were moved outwards along the short
  axis, the minimum-saddle-minimum set would merge into a single minimum,
  leaving a D system; the transition is known as a cusp catastrophe.
\item The middle row of Fig.~\ref{fig:splinput} shows three IQ
  systems: \sw{05}, \sw{42} and \sw{19}.  This type has a characteristic
  asymmetry, often with two images close together.  If the source
  were moved outwards, the minimum-saddle pair would merge and
  mutually cancel, leaving again a D system. This transition is known
  as a fold catastrophe.
\item The lower row of Fig.~\ref{fig:splinput} shows three SQ
  systems: \sw{09}, \sw{29} and \sw{02}. The failed model, \sw{57}, at top right
  may also belong to this category.  Here the images have a
  fairly symmetric arrangement with an arc and a counter-image on the
  other side, but the spaghetti diagram is completely asymmetric.  If
  the source were moved outwards along the long axis, the
  saddle-minimum-saddle set  would merge into a single saddle --- another
  form of cusp catastrophe.  SQ can be visually distinguished from LQ
  by the relative distances of the arc and the counter-image.  For
  SQ the arc is closer, for LQ the counter-image is closer.
\end{itemize}
CQ systems are often called `cross' or `Einstein-cross' systems; IQ
systems are sometimes called `folds', with `cusp' commonly used for
both SQ and LQ.  The labels `short-axis quad' and so on are not
standard in the literature, but the morphological classification they
express is familiar to experienced modellers.  Hence they can be
useful to researchers wishing to apply other modelling methods to the
same systems.
